---
title: "GSS challange"
author: "Diego Martinez"
date: "6/16/22"
output: html_document
---


```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2018.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

$H_0$: Attitude toward capital punishment is not related to race. Everyone has the same attitude and opinion towards capital punishment. 

$H_A$: Attitude toward capital punishment is dependent on race. Different races don't have the same attitude and opinion towards capital punishment. 


# Methods

Both race and cappun are categorical variables so the analysis technique is CAT~CAT; plot is a bar plot, summary is a cross-table, inference is a chi-squared test. We are going to use CAT vs CAT strategies since it'll be most effective comparing race vs their opinion on capital punishment. 

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results




<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->



## Now with The Most data (fewer variables gives more data points)

```{r}
dd2 <- mydata %>% group_by(race,cappun) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

From this plot, the white race is more strongly in favor of capital punishment compared to the other races. This is due to it's red bar plot (in favor) being much higher than it's bar that isn't in favor. Its in favor bar goes over 60%. That's much higher than the other two options presented. The Whites bar against the punishment is also the lowest out of all 3. In both aspects, White is most in favor. 

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->




## Now with More data

```{r}
table2 <- xtabs(~race + cappun, data=mydata)
rowPerc(table2)
colPerc(table2)
```


The white race has a higher rate in favor compared to any of the others. It's close to 68% in favor compared to the 46% in favor for African Americans. The white percentage in favor is 14% greater than the African American rate & is 12% greater than the other races rate. There's clearly a difference and the White continue to be most in favor. 


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

`


## More data version

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
#fisherXY(mydata$race,mydata$cappun)
```

The p-value for the chi-squared test was .00000000000002033. The p-value is very small so we reject the null hypotheses. We reject it because both p-values are way smaller than .05. The p-value for black vs other race is 0.02479 so regardless, both p-values given are different. 


```{r}
fisherXY(mydata$cappun,mydata$race)
```
 
The p-value for black vs other race is 0.02479 so regardless, both p-values given are different. The contingency table only further proves that there is a difference. 

# Discussion and Conclusion

Our question was whether or not different races had different opinions on capital punishment. We stated that both values are categorical so we worked around it. We used a bar plot, a table, and a chi-squared test. From the following, we learned that there is a difference. Both our p-values were below .05. The white race was way more in favor of capital punishment compared  to any other race by 14%. Thus, we rejected the null hypotheses due to there being way too much evidence against the claim that there was no correlation between race & their opinions. 

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
